---
title: "POLS 510 Student Project Assignment 1"
author: "Beatrice Magistro, Steven Otterman, and Lawrence Koster"
date: "10/10/2016"
output: html_document
---
  
  
  Our research project investigates which individual factors are more
likely to be associated with a Brexit vote. Specifically, we will examine whether
economic or cultural explanations are more likely to be associated
with an anti-EU stance. We will rely on two different datasets to test
our hypotheses and we will also eventually use official data from the
Referendum to check whether Brexit votes corresponded cross-regionally
with the dependent variables.

The two datasets we will be using are the Eurobarometer (EB) and the
European Social Survey (ESS). We will use the May 2015 Eurobarometer wave
and the 2014 European Social Survey wave. The unit of analysis is
individual respondents (potential voters from the UK). The dependent variable of interest from the EB dataset is: "Better future outside of the EU"", which is a discrete
variable that takes four possible values: Strongly agree (1), agree
(2), disagree (3), strongly disagree (4). The dependent variable of
interest from the ESS dataset is:  "Now thinking about the European
Union, some say European unification should go further. Others say it
has already gone too far. Using this card, what number on the scale
best describes your position?" A lower score (e.g. 1) means
"Unification already gone too far" and a higher score (e.g. 10)
"Unification go further".

The probability distribution that best describes our data are
multinomial distributions in both datasets. Two assumptions of a
multinomial distribution are that the trials it illustrates must be
independent and the probability of each possible result must be
constant. We do not think that either of these assumptions are violated
by that data. Furthermore, we believe that it would be appropriate to use MLE or ordinal logistic regression.

Our covariates of interest will be variables on the sujective and
objective feelings towards the state of the economy in the UK,
feelings towards immigration, and stances towards globalization, and we
will use socio-demographic controls such as age, education, gender, and
region as well.

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(rio)
library(dplyr)
library(magrittr)
library(ggplot2)
library(uwpols501)
library(tidyr)
library(broom)
library(uwpols501)
#library(haven)
#library(car)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
#EB <- import("EB.dta")
#save(EB, file = "EB.Rdata")
#EB <- read_dta("EB.dta")

load("EB.Rdata")

varnames <- c('unemplratecorrect','inflratecorrect', 'growthcorrect',
              'econ2correct'
)
for (i in varnames) {
  EB[[i]] <- ifelse(EB[[i]] %in% c(9),
                    NA, EB[[i]])
}

varnames <- c('finaexp', 'econexp', 'inflratecompared2013'
)
for (i in varnames) {
  EB[[i]] <- ifelse(EB[[i]] %in% c(4, 8, 9),
                    NA, EB[[i]])
}

varnames <- c('situationatecon', 'situationeuecon', 'finasituation',
              'feeleuimmigration', 'feelnoneuimmigration',
              'betteroutsideeu',
              'crisisreducepublicdebt','crisispublicdebtnopriority',
              'eusufficientpowers','crisisprivatesectorjobs',
              'crisispublicinvestment',
              'statspoldecisions', 'satisdmo', 'satisdeu',
              'citizen','stateintervenestoomuch',
              'toomuchtolerance', 'immigrantscontribute'
)
for (i in varnames) {
  EB[[i]] <- ifelse(EB[[i]] %in% c(5, 6, 7, 8, 9),
                    NA, EB[[i]])
}

varnames <- c('trustnp', 'trustgov',
              'trusteu','heardep',
              'heardec', 'heardecb', 'trustep', 'trustec', 'trustecb',
              'euknownomemb', 'euknowswimemb', 'euknowepelect', 'statstrust')
for (i in varnames) {
  EB[[i]] <- ifelse(EB[[i]] %in% c(3, 8, 9),
                    NA, EB[[i]])
}

varnames <- c('euimage', 'soclass', 'educrec5'
)
for (i in varnames) {
  EB[[i]] <- ifelse(EB[[i]] %in% c(6, 7, 8, 9),
                    NA, EB[[i]])
}


varnames <- c('unemplrate2014rec')
for (i in varnames) {
  EB[[i]] <- ifelse(EB[[i]] %in% c(8, 99),
                    NA, EB[[i]])
}

varnames <- c('inflrate2014rec', 'growth2014rec')
for (i in varnames) {
  EB[[i]] <- ifelse(EB[[i]] %in% c(9, 99),
                    NA, EB[[i]])
}

varnames <- c('unemplrate2014', 'inflrate2014', 'growth2014')
for (i in varnames) {
  EB[[i]] <- ifelse(EB[[i]] %in% c(99999, 999999),
                    NA, EB[[i]])
}

varnames <- c('married'
)
for (i in varnames) {
  EB[[i]] <- ifelse(EB[[i]] %in% c(15, 97),
                    NA, EB[[i]])
}

varnames <- c('educ'
)
for (i in varnames) {
  EB[[i]] <- ifelse(EB[[i]] %in% c(0, 97, 98, 99),
                    NA, EB[[i]])
}

varnames <- c('lrs', 'educrec11'
)
for (i in varnames) {
  EB[[i]] <- ifelse(EB[[i]] %in% c(97, 98),
                    NA, EB[[i]])
}

varnames <- c("regionBE",                  "regionFR",
              "regionNE",                   "regionDE",
              "regionIT",
              "regionLU",                   "regionDK",
              "regionIE",
              "regionUK",                   "regionGR",
              "regionES",
              "regionPT",                   "regionFI",
              "regionSE",
              "regionAU",                   "regionCY",
              "allregions",
              "regionCZ",                   "regionEE",
              "regionHU",
              "regionLV",                   "regionLT",
              "regionPL",
              "regionSK",                   "regionSI",
              "regionBG",
              "regionRO",                   "regionTR",
              "regionHR",
              "regionTCC",                  "regionMK",
              "regionRS",
              "regioneME",                  "regionAL"
)
for (i in varnames) {
  EB[[i]] <- ifelse(EB[[i]] %in% c(99),
                    NA, EB[[i]])
}

EB$age_gr <- cut(EB$age,
                 breaks = c(14, 25, 35, 45, 55, 65, 100),
                 labels = c("15-24", "25-34","35-44","45-54",
                            "55-64","65+"),
                 right = FALSE)

EB$country <- revalue(EB$country, c("DE-E" = "DE", "DE-W" = "DE", "GB-GBN" = "UK", 
                                    "GB-NIR" = "UK"))
UK <- filter(EB, country=='UK')
```
The first histrogram, below, is for the outcome variable from the Eurobarometer.

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(UK, aes(betteroutsideeu)) + geom_bar() + labs(x="Better future outside of the EU", y = "Count", title = "Eurobarometer 2015 Outcome Variable Histogram")
```
  
  
Next, the historgram for the outcome variable from the European Social Survey.
  
```{r echo=FALSE, warning=FALSE, message=FALSE}
load("ESS7GB.Rdata")
E1 <- ESSGB7
E1$euftf <- revalue(E1$euftf, c("Unification go further" = "10"))
E1 <- E1 %>% filter(!is.na(euftf))

ggplot(E1, aes(x = euftf)) + geom_bar() + labs(y = "Count", 
                    x = "European Unification Too Far or Not Far Enough (10 = Not Far Enough)",
                    title = "ESS 2014 Outcome Variable Histogram")
```
  
This final historgram groups the outcome variable from the European Social Survey into 2 groups: Less Unification (Strong) and More Unification (Strong). The first group contains all respondants who chose {1, 2, 3} as their responses while the More Unification (strong) group contains those who chose {8, 9, 10}. All other values {4, 5, 6, 7} as well as NAs were dropped.
  
```{r echo=FALSE, warning=FALSE, message=FALSE}
E3 <- ESSGB7
E3 <- E3 %>% filter(!is.na(euftf))
E3$euftf2[E3$euftf == 1 | E3$euftf == 2 | E3$euftf == 3] <- "LESS UNIFICATION (STRONG)"
E3$euftf2[E3$euftf == 8 | E3$euftf == 9 | E3$euftf == "Unification go further"] <- "MORE UNIFICATION (STRONG)"
ggplot(E3, aes(x = euftf2)) + geom_bar() + labs(y = "Count", 
                    x = "European Unification",
                    title = "ESS 2014 Outcome Variable Histogram (Strong)")
```